---
title: "variable_genes"
author: "EF"
date: "29 April 2016"
output: html_document
---
# start_section
  
```{r start_section}

# Time stamp
Sys.time()

# Clean up
rm(list=ls())
graphics.off()

# Set root working folder
library(knitr)
library(dplyr)
library(ggplot2)
library(stringr)

opts_knit$set(root.dir = "/scratch/medgen/users/eleanor/2015-16/R_scripts/")
setwd("/scratch/medgen/users/eleanor/2015-16/R_scripts/")

```
#Load original data
```{r original data}
exon <- read.table("ucsc_exonpos.txt", sep = "\t", quote="")
kgen <- read.table("nextera_exome_ip10_biallelic_nonsv_1k_vep_vcf.txt", sep = "\t", quote="")

```

#Edit data
```{r edit data}
library(stringr)
colnames(exon) <- c('CHR', 'Refgene','TYPE', 'START', 'STOP','N', 'STRAND', 'X', 'IDs')
exon2 <- data.frame(do.call('rbind', strsplit(as.character(exon$IDs),';',fixed=TRUE)))
exon3 <- data.frame(do.call('rbind', strsplit(as.character(exon2$X1),' ',fixed=TRUE)))
exon4 <- mutate(exon3, X2 = str_replace_all(X2, pattern='"', ''))
exon_clean <- cbind(exon, exon4)
exon.short <- select(exon_clean, CHR, TYPE, START, STOP, STRAND, X2)

rm(exon, exon2, exon3, exon4, exon_clean)
```

#Select cds and flip -ve strands
```{r test}
cds <- exon.short[exon.short$TYPE == "CDS",]
pos <- cds[cds$STRAND == "+",]
neg <- cds[cds$STRAND == "-",]
flip <- neg[order(neg$X2,neg$START, decreasing = TRUE),]
flipped_bind <- rbind(pos, flip)
rm(cds, exon.short, flip,neg, pos)
```

#Add exon number and length column
```{r add cols}
flipped_bind$exon <- ave(flipped_bind$X2 == flipped_bind$X2 , flipped_bind$X2, FUN=cumsum)

flipped_bind$length <- flipped_bind$STOP - flipped_bind$START
```

#Count number of variants in exon
```{r count variants}
count <- flipped_bind %>% rowwise() %>% 
  mutate(variants= nrow(kgen[kgen$POS >= START &  kgen$POS <= STOP,]))

#Add columns to data
```{r add col}
total_exons <- as.data.frame(count(exons_original.df, geneID))
shorted.df <- inner_join(short.df, total_exons, by="geneID")
shorted.df$exon_pos <- (shorted.df$exon/shorted.df$n)*100
colnames(shorted.df)[10] <- "total_exons"

```

#Set labels for variability
```{r set labels}
#percentage of variants above mean
mean_var <- median(shorted.df$variable)
(sum(shorted.df$variable > mean_var)/nrow(shorted.df))*100
summary <- as.table(summary(shorted.df$variable))
summary

#set values
low <- shorted.df[shorted.df$variable <= summary[2],]
low$score <- "low"
midlow <- shorted.df[shorted.df$variable > summary[2] & shorted.df$variable <= summary[3],]
midlow$score <- "mid_low"
midhigh <- shorted.df[shorted.df$variable > summary[3] & shorted.df$variable <= summary[5],]
midhigh$score <- "mid_high"
high <-shorted.df[shorted.df$variable > summary[5],]
high$score <- "high"

#merge rows and tidy
merge.scores.df <- bind_rows(low, midlow, midhigh, high)
colnames(merge.scores.df)[1] <- "CHROM"
clean.df <- mutate(merge.scores.df, CHROM=str_replace_all(CHROM, pattern="chr", ""))

#write csv
write.csv(clean.df, file= "variability_scores_file.csv")
```